A seamless, client-centric programming model for type safe web applications

### 2014-05-07 paper.tex
Add bootstrapping info to first example.

\titlebanner{-- ICFP '14 draft submission --}        % These are ignored unless
\preprintfooter{-- -- -- -- draft -- -- -- --}   % 'preprint' option specified.

@@

\begin{code}
main :: App Done
main = do
  greetings <- export helloServer

@@

The \lstinline!export! function takes an arbitrary function, provided that all
its arguments as well as its return value are serializable through the
\lstinline!Serialize! type class, and produces a typed identifier which
may be used to refer to the exported function. In this example, the type of
\lstinline!greetings! is \lstinline!Export (String -> Server ())!,
indicating that the identifier refers
to an exported function with a single \lstinline!String! argument and no return
value. Exported functions all live in the \lstinline!Server! monad.
This part of the program is executed on both the server and the client, albeit
with slightly different side effects, as described in section \ref{sec:impl}.

After the \lstinline!export! call, we enter the domain of client-exclusive code
with the application of \lstinline!runClient!. This function executes
computations in the \lstinline!Client! monad which is essentially an
IO monad with cooperative multitasking added on top, to mitigate the fact that
JavaScript has no native concurrency support. \lstinline!runClient! does not
return, and is the only function with a return type of \lstinline!App Done!,
which ensures that each \lstinline!App! computation contains exactly one client
computation.

In order to make an RPC call using an identifier obtained from
\lstinline!export!, we must supply it with an argument. This is done using the
\lstinline!<.>! operator. It might be interesting to note that its type,
\linebreak
\lstinline!Serialize a => Export (a -> b) -> a -> Export b!,
is very similar to the type of the \lstinline!<*>! operator over applicative
functors. This is not a coincidence; \lstinline!<.>! performs the same role
for the \lstinline!Export! type as \lstinline!<*>! performs for applicative
functors. The reason for using a separate operator for this instead of making
\lstinline!Export! an instance of \lstinline!Applicative! is that since
functions embedded in the \lstinline!Export! type exist only to be called over
a network, such functions must only be applied to arguments which can be
serialized and sent over a network connection. When an \lstinline!Export!
function is applied to an argument using \lstinline!<.>!, the argument is
serialized and stored inside the resulting \lstinline!Export! object, awaiting
dispatch. \lstinline!Export! values can thus be seen as explicit
representations of closures.

After applying the value obtained from the user to the exported function,
we apply the \lstinline!onServer! function to the result, which dispatches
the RPC call to the server. \lstinline!onServer! will then block until the
RPC call returns.

@@

main :: App Done
main = do
  greetings <- export helloServer

@@

@@

\begin{code}
fromRemote :: Remote a -> Server a
\end{code}

@@

  count <- export $ do
    r <- fromRemote remoteref
    liftIO $ atomicModifyIORef r (\v -> (v+1, v+1))

@@

\begin{code}
srvHello :: Remote RcptList -> Server ()
srvHello remoteRcpts = do
  recipients <- fromRemote remoteRcpts
  sid <- getSessionID
  liftIO $ do
    rcptMVar <- CC.newEmptyMVar
    CC.modifyMVar recipients $ \cs ->
      return ((sid, rcptMVar):cs ,())
\end{code}

@@

\begin{code}
srvSend :: Remote RcptList -> String -> Server ()
srvSend remoteRcpts message = do
    rcpts <- fromRemote remoteRcpts
    liftIO $ do
      recipients <- CC.readMVar rcpts
      mapM_ (CC.forkIO . deliver message) recipients
  where
    deliver msg (_,rcptMVar) = CC.putMVar rcptMVar msg
\end{code}

@@

\begin{code}
srvAwait :: Remote RcptList -> Server String
srvAwait remoteRcpts = do
  rcpts <- fromRemote remoteRcpts
  sid <- getSessionID
  liftIO $ do
    recipients <- CC.readMVar rcpts
    case lookup sid recipients of
      Just mv -> CC.takeMVar mv
      _       -> fail "Unregistered session!"
\end{code}

@@

Having implemented our three server operations, all that's left is to tie them
to the client. In this tying, we see our main advantage over the JavaScript
version in section \ref{sec:jsexample} in action: the \lstinline!export!
function builds a strongly typed bridge between the client and the server,
ensuring that any future enhancements to our chatbox program are made safely,
in one place, instead of being spread about throughout two disjoint code bases.

@@

  hello <- export $ srvHello recipients
  awaitMsg <- export $ srvAwait recipients
  sendMsg <- export $ srvSend recipients

@@

Notice that the \lstinline!recipients! list is passed to our three server
operations \emph{before} they are exported; since \lstinline!recipients! is
a mutable reference created on the server and inaccessible to client code,
it is not possible to pass it over the network as an RPC argument.
Even if it were possible, passing server-private state back and forth over the
network would be quite inappropriate due to privacy and security concerns.

The \lstinline!withElems! function is part of the Haste compiler's bundled DOM
manipulation library; it locates references to the DOM nodes with the given
identifiers and passes said references to a function.
In this case the variable \lstinline!log! will be bound to the node
with the identifier ``log'', and \lstinline!msgbox! will be bound to the node
identified by ``message''. These are the same DOM nodes that were referenced in
our original example, and refer to the chat log window and the text input
field respectively. After locating all the needed UI elements, the client
proceeds to register itself with the server's recipient list using the
\lstinline!hello! export.

@@

runClient    :: Client () -> App Done
liftServerIO :: IO a -> App (Remote a)
export       :: Exportable a
             => a -> App (Export a)

onServer     :: Export (Server a) -> Client a
(<.>)        :: Serialize a
             => Export (a -> b) -> a -> Export b

fromRemote   :: Remote a -> Server a
getSessionID :: Server SessionID
\end{code}
\caption{Types of the Haste.App core functions}
\label{lst:api}
\end{listingfloat}

\begin{table}
\renewcommand{\arraystretch}{1.5}
\begin{center}
\begin{tabular}{|r|l|}
\hline
Function & Purpose \\
\hline
\lstinline!runApp! & Begin execution of a Haste.App program. \\
\lstinline!runClient! & \parbox[t]{5cm}{Lift a single \lstinline!Client!
computation into the \lstinline!App! monad. Must be at the very end of
an \lstinline!App! computation, which is enforced by the type system.} \\
\lstinline!liftServerIO! & \parbox[t]{5cm}{Lift an IO computation into the
\lstinline!App! monad. The computation and its result are exclusive to the
server, as enforced by the type system, and are not observable on the client.} \\
\lstinline!export! & \parbox[t]{5cm}{Make a server side function available to
be called remotely by the client.} \\
\lstinline!onServer! & \parbox[t]{5cm}{Dispatch a remote call to the server and
wait for its completion. The result of the remote computation is returned on
the client after it completes.} \\
\lstinline!<.>! & \parbox[t]{5cm}{Apply an \lstinline!export!ed function to
a serializable argument.} \\
\lstinline!fromRemote! & \parbox[t]{5cm}{Extract the contents of a
\lstinline!Remote! value produced by \lstinline!liftServerIO! on the server.} \\
\lstinline!getSessionID! & \parbox[t]{5cm}{Get the unique identifier for
the current session. This is a pure convenience function, to relieve
programmers of the burden of session bookkeeping.} \\
\hline
\end{tabular}
\end{center}
\caption{Core functions of Haste.App}
\label{tbl:api}
\end{table}

@@

As we can see, \lstinline!App! is a simple state monad, with underlying IO
capabilities to allow server side computations to be forked from within it.
Its \lstinline!CallID! state element contains the identifier to be given to the
next exported function, and its other state element contains a mapping from
identifiers to exported functions.

@@

\begin{code}
data Remote a = Here a | There
data Export a = Export CallID [JSON]

liftServerIO :: IO a -> App (Remote a)
export :: Exportable a => a -> App (Export a)
\end{code}

@@

Its argument IO computation is executed, and the result is wrapped in the
\lstinline!Remote! constructor. The counterpart function, \lstinline!fromRemote!
simply unwraps the \lstinline!Here! constructor and returns the contained
value in the \lstinline!Server! monad.

\begin{code}
fromRemote :: Remote a -> Server a
fromRemote (Here x) = return x
\end{code}

\lstinline!export! is a more interesting function. Looking at its type, we see
that it accepts any \lstinline!Exportable! value for its argument.
\lstinline!Exportable! is defined as follows.

\begin{code}
class Exportable a where
  mkExported :: a -> ([JSON] -> Server JSON)

instance Serialize a => Exportable (Server a) where
  mkExported m = fmap toJSON m

instance (Serialize a, Exportable b) =>
         Exportable (a -> b) where
  mkExported f =
    \(x:xs) -> mkExported (f $ fromJSON x) xs
\end{code}

In essence, any function, over any number of arguments, which returns a
serializable value in the \lstinline!Server! monad can be exported. The
\lstinline!mkExported! function makes use of a well-known type class trick for
creating statically typed variadic functions, and works very much like the
\lstinline!printf! function of Haskell's standard library.\ \cite{printf}

The client implementation of \lstinline!export! is not very complex.

\begin{code}
export _ = App $ do
  (next_id, exports) <- get
  put (next_id+1, exports)
  return (Export next_id [])
\end{code}

As the exported function lives on the server, the client only needs an
identifier to be able to call on it. The exported function is thus ignored,
so that it can be optimized out of existence in the client version of the code.
The server version is only marginally more complex in that it updates the
identifier-function mapping table as well.

@@

\begin{code}
export f = App $ do
  (next_id, exports) <- get
  put (next_id+1, (next_id, mkExported f) : exports)
  return (Export next_id [])
\end{code}

The second argument of the \lstinline!Export! constructor is the list of
arguments to be sent to the export, to be updated as the exported function is
applied using the \lstinline!<.>! operator.

\begin{code}
(<.>) :: Serialize a
      => Export (a -> b) -> a -> Export b
(Export identifier args) <.> arg =
  Export identifier (toJSON arg : args)
\end{code}

@@

\begin{code}
onServer :: Serialize a
         => Export (Server a) -> Client a
onServer (Export identifier args) = do
  nonce <- get
  put (nonce + 1)
  mv <- createResultMVar nonce
  webSocketSend $
    toJSON (nonce, identifier, reverse args)  
  takeMVar mv
\end{code}

@@

The core of this interface consists of the \lstinline!ffi! function, which
allows the programmer to create a Haskell function from arbitrary JavaScript
code. This function exploits JavaScript's ability to parse and execute
arbitrary strings at run time using the \lstinline!eval! function, coupled with
the fact that functions in Haste and in JavaScript share the same
representation, to dynamically create a function object at runtime.
The \lstinline!ffi! function is typed using the same method as the
\lstinline!mkExported! function described in section \ref{sec:impl}.
When applied to one or more arguments instantiating the \lstinline!Marshal!
type class, the \lstinline!pack! function is applied to each argument,
marshalling them into their respective JavaScript representations, before they
are passed to the dynamically created function. When that function returns,
the inverse \lstinline!unpack! function is applied to its return value before
it is passed back into the Haskell world.

@@

Several other approaches to the seamless client-server interaction exist. In
general, these attacks tend to be of the ``all or nothing'' variety,
introducing new languages or otherwise requiring custom full stack solutions.
In contrast, our solution can be implemented entirely as a library and is
portable to any pair of compilers supporting typed monadic programming.
Moreover, Haste.App has a quite simple and controlled programming model with a
clearly defined controller, which stands in contrast to most related work which
embraces a more flexible but also more complex programming model.

@@

\paragraph{Inspiration and alternatives to \lstinline!export!}
\label{sec:cloudhaskell}
One crucial aspect of implementing cross-network function calls is the issue of
data representation: the client side of things must be able to obtain some
representation of any function it may want to call on the server.

In our solution, this representation is obtained through the use of the \lstinline!export! function, which when executed on the server pairs a function with a
unique identifier and on the client, and when executed on the client returns
said identifier so that the client may now refer to the function. While this
has the advantage of being simple to implement, one major drawback of this
method is that all functions must be explicitly exported in the \lstinline!App!
monad prior to being called over the network.

@@

The astute reader may notice that this is exactly what the \lstinline!export!
function does as well, the difference being that \lstinline!export! links the
identifier to the value it represents on the type level, making it impossible
to call non-existent remote functions and break the program's type safety in
other ways.

>>>>

\titlebanner{-- Haskell Symposium '14 draft submission --}        % These are ignored unless
\preprintfooter{-- -- -- -- draft -- -- -- --}   % 'preprint' option specified.

@@

\begin{code}
main :: App Done
main = do
  greetings <- remote helloServer

@@

The \lstinline!remote! function takes an arbitrary function, provided that all
its arguments as well as its return value are serializable through the
\lstinline!Serialize! type class, and produces a typed identifier which
may be used to refer to the remote function. In this example, the type of
\lstinline!greetings! is \lstinline!Remote (String -> Server ())!,
indicating that the identifier refers
to a remote function with a single \lstinline!String! argument and no return
value. Remote functions all live in the \lstinline!Server! monad.
This part of the program is executed on both the server and the client, albeit
with slightly different side effects, as described in section \ref{sec:impl}.

After the \lstinline!remote! call, we enter the domain of client-exclusive code
with the application of \lstinline!runClient!. This function executes
computations in the \lstinline!Client! monad which is essentially an
IO monad with cooperative multitasking added on top, to mitigate the fact that
JavaScript has no native concurrency support. \lstinline!runClient! does not
return, and is the only function with a return type of \lstinline!App Done!,
which ensures that each \lstinline!App! computation contains exactly one client
computation.

In order to make an RPC call using an identifier obtained from
\lstinline!remote!, we must supply it with an argument. This is done using the
\lstinline!<.>! operator. It might be interesting to note that its type,
\linebreak
\lstinline!Serialize a => Remote (a -> b) -> a -> Remote b!,
is very similar to the type of the \lstinline!<*>! operator over applicative
functors. This is not a coincidence; \lstinline!<.>! performs the same role
for the \lstinline!Remote! type as \lstinline!<*>! performs for applicative
functors. The reason for using a separate operator for this instead of making
\lstinline!Remote! an instance of \lstinline!Applicative! is that since
functions embedded in the \lstinline!Remote! type exist only to be called over
a network, such functions must only be applied to arguments which can be
serialized and sent over a network connection. When a \lstinline!Remote!
function is applied to an argument using \lstinline!<.>!, the argument is
serialized and stored inside the resulting \lstinline!Remote! object, awaiting
dispatch. \lstinline!Remote! computations can thus be seen as explicit
representations of closures.

After applying the value obtained from the user to the remote function,
we apply the \lstinline!onServer! function to the result, which dispatches
the RPC call to the server. \lstinline!onServer! will then block until the
RPC call returns.

@@

main :: App Done
main = do
  greetings <- remote helloServer

@@

To run this example, an address and a port must be provided so that the client
knows which server to contact. There are several ways of doing this:
using the GHC plugin system, through Template Haskell or by slightly altering
how program entry points are treated in a compiler or wrapper script, to name
a few. A non-intrusive method when using the GHC/Haste compiler pair would
be to pass add \lstinline!-main-is setup! to both compilers' command line
and add the \lstinline!setup! function in figure \ref{lst:setup} to the source
code. This will instruct the server binary to listen on the port 1111 when
started, and the client to attempt contact with that port on the local machine.

The exact mechanism chosen to provide the host and port are implementation
specific, and will in the interest of clarity not be discussed further.

\begin{listingfloat}
\begin{code}
setup :: IO ()
setup =
  runApp (mkConfig "ws://localhost:1111" 1111) main
\end{code}
\caption{Bootstrapping a Haste.App program}
\label{lst:setup}
\end{listingfloat}

@@

\begin{code}
use :: Remote a -> Server a
\end{code}

@@

  count <- remote $ do
    r <- use remoteref
    liftIO $ atomicModifyIORef r (\v -> (v+1, v+1))

@@

\begin{code}
srvHello :: Remote RcptList -> Server ()
srvHello remoteRcpts = do
  recipients <- use remoteRcpts
  sid <- getSessionID
  liftIO $ do
    rcptMVar <- CC.newEmptyMVar
    CC.modifyMVar recipients $ \cs ->
      return ((sid, rcptMVar):cs ,())
\end{code}

@@

\begin{code}
srvSend :: Remote RcptList -> String -> Server ()
srvSend remoteRcpts message = do
    rcpts <- use remoteRcpts
    liftIO $ do
      recipients <- CC.readMVar rcpts
      mapM_ (CC.forkIO . deliver message) recipients
  where
    deliver msg (_,rcptMVar) = CC.putMVar rcptMVar msg
\end{code}

@@

\begin{code}
srvAwait :: Remote RcptList -> Server String
srvAwait remoteRcpts = do
  rcpts <- use remoteRcpts
  sid <- getSessionID
  liftIO $ do
    recipients <- CC.readMVar rcpts
    case lookup sid recipients of
      Just mv -> CC.takeMVar mv
      _       -> fail "Unregistered session!"
\end{code}

@@

Having implemented our three server operations, all that's left is to tie them
to the client. In this tying, we see our main advantage over the JavaScript
version in section \ref{sec:jsexample} in action: the \lstinline!remote!
function builds a strongly typed bridge between the client and the server,
ensuring that any future enhancements to our chatbox program are made safely,
in one place, instead of being spread about throughout two disjoint code bases.

@@

  hello <- remote $ srvHello recipients
  awaitMsg <- remote $ srvAwait recipients
  sendMsg <- remote $ srvSend recipients

@@

Notice that the \lstinline!recipients! list is passed to our three server
operations \emph{before} they are imported; since \lstinline!recipients! is
a mutable reference created on the server and inaccessible to client code,
it is not possible to pass it over the network as an RPC argument.
Even if it were possible, passing server-private state back and forth over the
network would be quite inappropriate due to privacy and security concerns.

The \lstinline!withElems! function is part of the Haste compiler's bundled DOM
manipulation library; it locates references to the DOM nodes with the given
identifiers and passes said references to a function.
In this case the variable \lstinline!log! will be bound to the node
with the identifier ``log'', and \lstinline!msgbox! will be bound to the node
identified by ``message''. These are the same DOM nodes that were referenced in
our original example, and refer to the chat log window and the text input
field respectively. After locating all the needed UI elements, the client
proceeds to register itself with the server's recipient list using the
\lstinline!hello! remote computation.

@@

runClient    :: Client () -> App Done
liftServerIO :: IO a -> App (RemoteValue a)
remote       :: Remotable a
             => a -> App (Remote a)

onServer     :: Remote (Server a) -> Client a
(<.>)        :: Serialize a
             => Remote (a -> b) -> a -> Remote b

use          :: RemoteValue a -> Server a
getSessionID :: Server SessionID
\end{code}
\caption{Types of the Haste.App core functions}
\label{lst:api}
\end{listingfloat}

\begin{table}
\renewcommand{\arraystretch}{1.5}
\begin{center}
\begin{tabular}{|r|l|}
\hline
Function & Purpose \\
\hline
\lstinline!runApp! & Begin execution of a Haste.App program. \\
\lstinline!runClient! & \parbox[t]{5cm}{Lift a single \lstinline!Client!
computation into the \lstinline!App! monad. Must be at the very end of
an \lstinline!App! computation, which is enforced by the type system.} \\
\lstinline!liftServerIO! & \parbox[t]{5cm}{Lift an IO computation into the
\lstinline!App! monad. The computation and its result are exclusive to the
server, as enforced by the type system, and are not observable on the client.} \\
\lstinline!remote! & \parbox[t]{5cm}{Make a server side function available to
be called remotely by the client.} \\
\lstinline!onServer! & \parbox[t]{5cm}{Dispatch a remote call to the server and
wait for its completion. The result of the remote computation is returned on
the client after it completes.} \\
\lstinline!<.>! & \parbox[t]{5cm}{Apply an \lstinline!remote! function to
a serializable argument.} \\
\lstinline!use! & \parbox[t]{5cm}{Extract the contents of a
\lstinline!RemoteValue! produced by \lstinline!liftServerIO! on the server.} \\
\lstinline!getSessionID! & \parbox[t]{5cm}{Get the unique identifier for
the current session. This is a pure convenience function, to relieve
programmers of the burden of session bookkeeping.} \\
\hline
\end{tabular}
\end{center}
\caption{Core functions of Haste.App}
\label{tbl:api}
\end{table}

@@

As we can see, \lstinline!App! is a simple state monad, with underlying IO
capabilities to allow server side computations to be forked from within it.
Its \lstinline!CallID! state element contains the identifier to be given to the
next remote function, and its other state element contains a mapping from
identifiers to remote functions.

@@

\begin{code}
data RemoteValue a = Here a | There
data Remote a = Remote CallID [JSON]

liftServerIO :: IO a -> App (RemoteValue a)
remote :: Remotable a => a -> App (Remote a)
\end{code}

@@

Its argument IO computation is executed, and the result is wrapped in the
\lstinline!Remote! constructor. The counterpart function, \lstinline!use!
simply unwraps the \lstinline!Here! constructor and returns the contained
value in the \lstinline!Server! monad.

\begin{code}
use :: Remote a -> Server a
use (Here x) = return x
\end{code}

\lstinline!remote! is a more interesting function. Looking at its type, we see
that it accepts any \lstinline!Remotable! value for its argument.
\lstinline!Remotable! is defined as follows.

\begin{code}
class Remotable a where
  mkRemote :: a -> ([JSON] -> Server JSON)

instance Serialize a => Remotable (Server a) where
  mkRemote m = fmap toJSON m

instance (Serialize a, Remotable b) =>
         Remotable (a -> b) where
  mkRemote f =
    \(x:xs) -> mkRemote (f $ fromJSON x) xs
\end{code}

In essence, any function, over any number of arguments, which returns a
serializable value in the \lstinline!Server! monad can be imported. The
\lstinline!mkRemote! function makes use of a well-known type class trick for
creating statically typed variadic functions, and works very much like the
\lstinline!printf! function of Haskell's standard library.\ \cite{printf}

The client implementation of \lstinline!remote! is not very complex.

\begin{code}
remote _ = App $ do
  (next_id, remotes) <- get
  put (next_id+1, remotes)
  return (Remote next_id [])
\end{code}

As the remote function lives on the server, the client only needs an
identifier to be able to call on it. The remote function is thus ignored,
so that it can be optimized out of existence in the client version of the code.
The server version is only marginally more complex in that it updates the
identifier-function mapping table as well.

@@

\begin{code}
remote f = App $ do
  (next_id, remotes) <- get
  put (next_id+1, (next_id, mkRemote f) : remotes)
  return (Remote next_id [])
\end{code}

The second argument of the \lstinline!Remote! constructor is the list of
arguments to be sent to the function, to be updated as the remote function is
applied using the \lstinline!<.>! operator.

\begin{code}
(<.>) :: Serialize a
      => Remote (a -> b) -> a -> Remote b
(Remote identifier args) <.> arg =
  Remote identifier (toJSON arg : args)
\end{code}

@@

\begin{code}
onServer :: Serialize a
         => Remote (Server a) -> Client a
onServer (Remote identifier args) = do
  nonce <- get
  put (nonce + 1)
  mv <- createResultMVar nonce
  webSocketSend $
    toJSON (nonce, identifier, reverse args)  
  takeMVar mv
\end{code}

@@

The core of this interface consists of the \lstinline!ffi! function, which
allows the programmer to create a Haskell function from arbitrary JavaScript
code. This function exploits JavaScript's ability to parse and execute
arbitrary strings at run time using the \lstinline!eval! function, coupled with
the fact that functions in Haste and in JavaScript share the same
representation, to dynamically create a function object at runtime.
The \lstinline!ffi! function is typed using the same method as the
\lstinline!mkRemote! function described in section \ref{sec:impl}.
When applied to one or more arguments instantiating the \lstinline!Marshal!
type class, the \lstinline!pack! function is applied to each argument,
marshalling them into their respective JavaScript representations, before they
are passed to the dynamically created function. When that function returns,
the inverse \lstinline!unpack! function is applied to its return value before
it is passed back into the Haskell world.

@@

Several other approaches to seamless client-server interaction exist. In
general, these proposed solutions tend to be of the ``all or nothing'' variety,
introducing new languages or otherwise requiring custom full stack solutions.
In contrast, our solution can be implemented entirely as a library and is
portable to any pair of compilers supporting typed monadic programming.
Moreover, Haste.App has a quite simple and controlled programming model with a
clearly defined controller, which stands in contrast to most related work which
embraces a more flexible but also more complex programming model.

@@

\paragraph{Inspiration and alternatives to \lstinline!remote!}
\label{sec:cloudhaskell}
One crucial aspect of implementing cross-network function calls is the issue of
data representation: the client side of things must be able to obtain some
representation of any function it may want to call on the server.

In our solution, this representation is obtained through the use of the \lstinline!remote! function, which when executed on the server pairs a function with a
unique identifier and on the client, and when executed on the client returns
said identifier so that the client may now refer to the function. While this
has the advantage of being simple to implement, one major drawback of this
method is that all functions must be explicitly imported in the \lstinline!App!
monad prior to being called over the network.

@@

The astute reader may notice that this is exactly what the \lstinline!remote!
function does as well, the difference being that \lstinline!remote! links the
identifier to the value it represents on the type level, making it impossible
to call non-existent remote functions and break the program's type safety in
other ways.
